This is the title [Article v\versionnumber]

### 2022-10-14 manuscript/manuscript.tex

\begin{equation}
g_\textit{vox} ( \textbf{r}, \omega ) \approx g(\textbf{r}) g(\omega)
\end{equation}

@@

\section{Recent developments}
In recent years, several updates to the original GIST implementation have been published.
The basic functionality of the program, however, has not been changed.
In this chapter, we will shortly present each of those updates.
\subsection{GPU implementation}
In Reference~\cite{Kraml2020}, a GPU implementation of the energy calculation was implemented, which typically increases the speed by 1--2 orders of magnitude.

If \software{cpptraj} is compiled with GPU support, the energy calculation uses the GPU automatically, without any additional input.
However, PME-GIST is not supported on the GPU.

\subsection{PME implementation}
In Reference~\cite{Chen2021}, a PME implementation of the GIST energy was presented.
For typical systems, this implementation is slightly slower than GPU-GIST, but much faster than the original CPU code.
Furthermore, PME-GIST offers the best agreement between GIST energies and those observed in an MD simulation.

To run a PME-GIST calculation, the \inlinecode{pme} flag to the \inlinecode{gist} action in \software{cpptraj} can be used.
The output will contain two additional columns: \inlinecode{PME\_dens} and \inlinecode{PME\_norm}.
They contain the potential of solute and solvent evaluated at the solvent positions and divided by two.
Furthermore, the \inlinecode{Eww} and \inlinecode{Esw} columns are also computed using PME, and are reported as usual.
For typical use cases, it is recommended to use \inlinecode{Eww} and \inlinecode{Esw} over the \inlinecode{PME} column.

\subsection{GIST with non-water solvents}
In References~\cite{Kraml2020,Kamenik2020-gist-macrocycles,Waibl2022-gist-solvents}, an extension of GIST towards solvents other than water is described.

To run a GIST calculation in a solvent other than water, use the \inlinecode{solute} flag together with a \software{cpptraj} selection mask.
Everything that is not in the \inlinecode{solute} mask is treated as solvent.
If there are multiple solvent species, use the \inlinecode{solventmols} option as described in Section~\ref{sec-salt-water}.
The code will automatically choose three atoms which are used to define the solvent orientation, and print this selection to the output.

Sometimes, \software{cpptraj} does not automatically choose the optimal atoms.
In methanol, for instance, one can either incorporate the C--H bonds or the O--H bond in the selection.
The latter is probably more relevant since it can incorporate hydrogen bonding effects.
Assuming that the alcohol hydrogen is called H1, this could be specified as follows:

\begin{lstlisting}[style=cpptraj]
gist gridcntr <x> <y> <z> \
griddim <Nx> <Ny> <Nz> gridspacn <val> \
solute ^1 rigidatoms O1 H1 C1
\end{lstlisting}

Note that the central atom of the \inlinecode{rigidatoms} selection goes first.

\subsection{GIST with salt-water mixtures}
\label{sec-salt-water}
In \cite{Waibl2021-gist-salt}, an extension of GIST was presented that can use salt-water mixtures as a solvent.
This allows treating salting-out effects, although it was shown that the salting-out coefficient is over-predicted by GIST unless higher-order entropies are incorporated.
Generally, GIST should be able to treat arbitrary solvent mixtures as long as each molecule is sufficiently rigid.

@@

\begin{figure}
	\centering
	\includegraphics[width=1.0\linewidth]{streptavidin_bp.png}
	\caption{Volume of strepatividin defined around 3.5 A of biotin. }\label{streptavidin_sasa}
\end{figure}

>>>>

\begin{equation}
g_\textit{vox} \left( \textbf{r}, \omega \right) \approx g_\textit{vox}(\textbf{r}) g_\textit{vox}(\omega)
\end{equation}

@@

\section{Recent developments}
\subsection{GIST with non-water solvents}
\subsection{GIST with salt-water mixtures}
In \cite{Waibl2021-gist-salt}, an extension of GIST was presented that can use salt-water mixtures as a solvent.
This allows treating salting-out effects, although it was shown that the salting-out coefficient is over-predicted by GIST unless higher-order entropies are incorporated.
Generally, GIST should be able to treat arbitrary solvent mixtures as long as each molecule is sufficiently rigid.

@@

\begin{figure}
	\centering
	\includegraphics[width=1.0\linewidth]{streptavidin_bp_close.png}
	\caption{Volume of strepatividin defined around 3.5 A of biotin. }\label{streptavidin_bp_close}
\end{figure}
